Drug sensitivity prediction with normal inverse Gaussian shrinkage informed by external data

### 2020-04-16 docs/supplement1.Rnw
commit before pull

tab3 <- sapply(1:ncol(tabm), function(s) {
  id <- which(tabm[, s]==min(tabm[, s], na.rm=TRUE))
  out <- paste0(round(tabm[, s], 3), " (", round(tabsd[, s], 3), ")")
  out[id] <- paste0("\\textbf{", out[id], "}")
  out})
colnames(tab3) <- c("$\\text{EMSE}_{\\text{total}}$", 
                    "$\\text{EMSE}_{\\text{top}}$", 
                    "$\\text{EMSE}_{\\text{bottom}}$", "$\\text{PMSE}$")     
rownames(tab3) <- c("NIG$_{\\text{d}}^-$", 
                     "NIG$_{\\text{d}}$", "ridge", "lasso", "mxtune")
@	
	Tables \ref{tab:simulations_gdsc_tab1}-\ref{tab:simulation_gdsc_tab3} 
	display averaged estimation mean squared error (EMSE), and prediction mean
	squared error (PMSE), calculated on independent test data, for simulation 
	Scenarios 1-4. PMSE is calculated as in Section \ref{md-sec:gdsc} of the MD, 
	while EMSE is calculated as: $\text{EMSE} = D^{-1} p^{-1} \sum_{d=1}^D 
  \sum_{j=1}^p (\beta_{jd} - \hat{\beta}_{jd})^2$, with
  $\hat{\beta}_{jd}$ the estimator for $\beta_{jd}$. In the NIG
  models, that provide full posteriors, the posterior mean 
  $\E(\beta_{jd} | \mathbf{y}_d)$ is used as point estimate.
	EMSE is further split into the contribution of the
	bottom 90\% of the $\beta_{jd}$ in terms of size (in absolute value) and the 
	top 10\% of the $\beta_{jd}$ in size.

%   
%   
%   $\lambda_{\text{feat}}$ is overestimated with median estimate 
% 	\Sexpr{round(est.lambdaf[2], 2)} (the
% 	true value is 1). The overestimated $\lambda_{\text{feat}}$ translates 
% 	into underestimated $\gamma_{jd}^2$ variances. 
% 	
% 	
% 	
<<simulations_gdsc_tab1>>=
kableExtra::kable_styling(knitr::kable(
  tab1, align="r",
  caption="Mean EMSE (split for bottom 10\\% and top 90\\% of $|\\hat{\\beta}_{jd}|$) and PMSE for simulation Scenario 1, estimated on test data (lowest in bold).", 
  format="latex", booktabs=TRUE, escape=FALSE), 
  latex_options=c("HOLD_position"))
@

<<simulations_gdsc_tab2>>=
kableExtra::kable_styling(knitr::kable(
  tab2, align="r",
  caption="Mean EMSE (split for bottom 10\\% and top 90\\% of $|\\hat{\\beta}_{jd}|$) and PMSE for simulation Scenario 2, estimated on test data (lowest in bold).", 
  format="latex", booktabs=TRUE, escape=FALSE), 
  latex_options=c("HOLD_position"))
@

<<simulations_gdsc_tab3>>=
kableExtra::kable_styling(knitr::kable(
  tab3, align="r",
  caption="Mean EMSE (split for bottom 10\\% and top 90\\% of $|\\hat{\\beta}_{jd}|$) and PMSE for simulation Scenario 2, estimated on test data (lowest in bold).", 
  format="latex", booktabs=TRUE, escape=FALSE), 
  latex_options=c("HOLD_position"))
@

<<simulations_gdsc_pmse4, fig.cap="Simulation results for Scenario 4: Average PMSE against the proportion of permuted rows in the external features.", out.width="100%", fig.asp=1>>= 
@

	Focussing on EMSE, we see that the NIG$_{\text{f}}$ model,
	that includes the external covariates, performs best, both in terms of 
	estimation and prediction. The lower MSEs confirm that the model that
	includes the external data (NIG$_{\text{f}}$) learns the underlying structure 
	in the data better than model NIG$_{\text{f}}^-$ that does not include the
	external data. Furthermore, we see that estimation of the larger $\beta_{jd}$ 
	is where ridge loses the most compared to the NIG and lasso models. 
	The heavier tails of the NIG and lasso priors allow for larger $\beta_{jd}$
	estimates and thus lower EMSE for these $\beta_{jd}$.

%   $\lambda_{\text{feat}}$ is overestimated with median estimate
% 	\Sexpr{round(est.lambdaf[1], 2)} (the true value is one) and thus 
% 	again underestimates the 
% 	prior $\gamma_{jd}^2$ variance.
% 	
% 	Median $\lambda_{\text{drug}}=\Sexpr{round(est.lambdad[2], 2)}$ is
% 	overestimated compared to true value one. 
% 	Scale $\lambda_{\text{drug}}$ is underestimated with median estimate
% 	\Sexpr{round(est.lambdad[1], 2)} compared to the true value one.
% 	
%   Median performance measures for simulation Scenario 2 are displayed in Table
% 	\ref{tab:simulations_gdsc_tab2}. Again, we see that NIG$_{\text{d}}$ performs
% 	best in terms of MSE. However, NIG$_{\text{d}}^-$ performance is the same 
% 	here, indicating that the external covariates do not add much here.
% 	Similarly as in Scenario 1, most of the estimation performance gain over ridge
% 	is through the top 10\% of the $\beta_{jd}$.
% 	
%   
%   Median $\lambda_{\text{feat}}$ and $\lambda_{\text{drug}}$ are
% 	are over- and underestimated with \Sexpr{round(est.lambdaf[2], 2)}
% 	and \Sexpr{round(est.lambdad[2], 2)}, respectively, compared to their true
% 	value 1. 
% 	
%   NIG$_{\text{f}+\text{d}}^-$ also over- and
% 	underestimates 
% 	$\lambda_{\text{feat}}$ and $\lambda_{\text{drug}}$ with median estimates
% 	\Sexpr{round(est.lambdaf[1], 2)}
% 	and \Sexpr{round(est.lambdad[1], 2)}, respectively. 
% 	
%   Table 
% 	\ref{tab:simulations_gdsc_tab3}, with median performance measures, shows
% 	that NIG$_{\text{f}+\text{d}}$ outperforms the other models in terms of
% 	prediction and estimation MSE, with again most of the edge over ridge in the
% 	top 10\% of the $\beta_{jd}$.

@@

<<analysis_gdsc_res2, fig.cap="Simulation results for Scenario 4: Median PMSE against the proportion of permuted rows in the external features.", out.width="50%", fig.asp=1>>= 
@

>>>>

tab3 <- sapply(1:ncol(tabm), function(s) {
  id <- which(tabm[, s]==min(tabm[, s], na.rm=TRUE))
  out <- paste0(round(tabm[, s], 3), " (", round(tabsd[, s], 3), ")")
  out[id] <- paste0("\\textbf{", out[id], "}")
  out})
colnames(tab3) <- c("$\\text{EMSE}_{\\text{total}}$", 
                    "$\\text{EMSE}_{\\text{top}}$", 
                    "$\\text{EMSE}_{\\text{bottom}}$", "$\\text{PMSE}$")     
rownames(tab3) <- c("NIG$_{\\text{f+d}}^-$", 
                    "NIG$_{\\text{f+d}}$", "ridge", "lasso", "mxtune")
@	
	Tables \ref{tab:simulations_gdsc_tab1}-\ref{tab:simulations_gdsc_tab3} 
	display averaged estimation mean squared error (EMSE), and prediction mean
	squared error (PMSE), calculated on independent test data, for simulation 
	Scenarios 1-4. PMSE is calculated as in Section \ref{md-sec:gdsc} of the MD, 
	while EMSE is calculated as: $\text{EMSE} = D^{-1} p^{-1} \sum_{d=1}^D 
  \sum_{j=1}^p (\beta_{jd} - \hat{\beta}_{jd})^2$, with
  $\hat{\beta}_{jd}$ the estimator for $\beta_{jd}$. In the NIG
  models, that provide full posteriors, the posterior mean 
  $\E(\beta_{jd} | \mathbf{y}_d)$ is used as point estimate.
	EMSE is further split into the contribution of the
	bottom 10\% of the $\beta_{jd}$ in terms of size (in absolute value) and the 
	top 10\% of the $\beta_{jd}$ in size.

	Focussing on estimation, we see that the NIG$_{\text{f}}$ model,
	that includes the external covariates, performs best, both in terms of 
	estimation and prediction. The lower MSEs confirm that the model that
	includes the external data (NIG$_{\text{f}}$) learns the underlying structure 
	in the data better than model NIG$_{\text{f}}^-$ that does not include the
	external data. Furthermore, we see that estimation of the larger $\beta_{jd}$ 
	is where ridge loses the most compared to the NIG and lasso models. 
	The heavier tails of the NIG and lasso priors allow for larger $\beta_{jd}$
	estimates and thus lower EMSE for these $\beta_{jd}$.

<<simulations_gdsc_tab1>>=
kableExtra::kable_styling(knitr::kable(
  tab1, align="r",
  caption="Mean EMSE (separated for bottom and top 10\\% of $|\\hat{\\beta}_{jd}|$) and PMSE for simulation Scenario 1, estimated on test data (lowest in bold).", 
  format="latex", booktabs=TRUE, escape=FALSE), 
  latex_options=c("HOLD_position"))
@

<<simulations_gdsc_tab2>>=
kableExtra::kable_styling(knitr::kable(
  tab2, align="r",
  caption="Mean EMSE (separated for bottom and top 10\\% of $|\\hat{\\beta}_{jd}|$) and PMSE for simulation Scenario 2, estimated on test data (lowest in bold).", 
  format="latex", booktabs=TRUE, escape=FALSE), 
  latex_options=c("HOLD_position"))
@

<<simulations_gdsc_tab3>>=
kableExtra::kable_styling(knitr::kable(
  tab3, align="r",
  caption="Mean EMSE (separated for bottom and top 10\\% of $|\\hat{\\beta}_{jd}|$) and PMSE for simulation Scenario 3, estimated on test data (lowest in bold).", 
  format="latex", booktabs=TRUE, escape=FALSE), 
  latex_options=c("HOLD_position"))
@

<<simulations_gdsc_pmse4, fig.cap="(a) Mean EMSE, separated for (b) bottom and (c) top 10\\% of $|\\hat{\\beta}_{jd}|$, and (d) PMSE for simulation Scenario 3 versus the proportion of permuted  rows in the external data, estimated on test data.", out.width="100%", fig.asp=1>>= 
@

@@

<<analysis_gdsc_res2, fig.cap="Simulation results for Scenario 4: Median PMSE against the proportion of permuted rows in the external features.", out.width="100%", fig.asp=1>>= 
@
